---
title: "Seattle_Airbnb"
author: "Patrick McKee"
date: "July 30, 2020"
output: html_document
---

##Data Import

```{r setup, include=FALSE}
###Read in data
library(readr)
data = read_csv("Copy of seattle_01.csv")

```

## EDA of Listing Data

```{r EDA}
##Structure of Airbnb data
str(data)
summary(data)
####Remove unneeded columns
data$rate_type <- NULL
data$currency <- NULL
data$X1 <- NULL
data$room_id <- NULL
data$host_id <- NULL
data$address <- NULL
data$latitude <- NULL
data$longitude <- NULL
data$name <- NULL
data$location <- NULL
data$last_modified <- NULL

###final str of data
str(data)

```

## Check and Replace for Missing Values

```{r Missing Value Check, echo=FALSE}
###Missing value check in dataframe
sum(is.na(data))
###Display how many missing values in each column
colSums(is.na(data))
```

Overall_satisfaction column is missing 1473 values, roughly 20% less than our total amount of Airbnb listings. Need to impute values to fill out missing 20%.

```{r Bathroom missing values}
###Bathroom missing values
###Replace missing bathroom values with median
data$bathrooms[(is.na(data$bathrooms))] <- median(data$bathrooms,na.rm =T)
###let us impute all the the missing value with mean
data$overall_satisfaction[is.na(data$overall_satisfaction)] <-  mean(data$overall_satisfaction,na.rm =T)

```

Checking for correlation between price and overall_satisfaction

```{r Overall_Satisfaction missing values}
####Overall Satisfaction missing values
table(data$overall_satisfaction)
####check the correlation betwwen the price and coorelation
data_complete <- data[complete.cases(data),]
cor(data_complete$price,data_complete$overall_satisfaction)


```

Room type summary and distribution

```{r Room Type Factor Configuration}
####Room Type structure
str(data$room_type)
###distribution of room type
table(data$room_type)

##summary of data
summary(data)

###converting review numbers into categorial column
data$reviews <- ifelse(data$reviews > 100, "Very high", ifelse(data$reviews > 64, "high",ifelse(data$reviews > 47, "med",ifelse(data$reviews > 21 ,"low", 'very low'))))
###distrubution
table(data$reviews)
##str
str(data$reviews)
###converting charcter into factor
data$reviews <- as.factor(data$reviews)


```

##Outlier Checks

```{r Accomodates Checks}

##Accommodates Outlier check
boxplot(data$accommodates)
###anything abbove 10 is outlier in accomodates
###putting > 10  to 10
data$accommodates[data$accommodates > 10] <- 10
boxplot(data$accommodates, xlab = "After Ouliers Removed", ylab ="accommodates")

```

```{r Price Outlier}
##Price Outlier check
boxplot(data$price)
###anything abbove 3000 is outlier in price
###putting > 3000  to 3000
data$price[data$price > 3000] <- 3000
boxplot(data$price, xlab = "After Ouliers Removed", ylab ="price")
```

## Creating Price Classes

```{r Price category Class}
##converting price into classes,
##1. low  2.median 3.high
summary(data$price)
data$price_update <- ifelse(data$price <= 88 , "low", ifelse((data$price > 88 & data$price < 125), "median", ifelse(data$price >= 125, "high","others")))

#New price class distribution & structure
table(data$price_update)
str(data$price_update)
###converting price in factor
data$price_update <- as.factor(data$price_update)
###removing price column
data$price <- NULL
```

Created classes for prices in Seattle market. Anything equal to or less than \$88 is "Low", greater than \$88 and less than \$125 is "Median", and anything greather than \$125 is "High". 

##Train and Test Sets

```{r Train and Test}
###Training and Test Set
library(caret)
set.seed(1234)
intrain <- createDataPartition(y = data$price_update, p = 0.8, list = FALSE)
training <- data[intrain,]
testing <- data[-intrain,]

###seeing the distrubution of data
table(training$price_update)

table(testing$price_update)
```

##Decision Tree
```{r Decision Tree and rpart}
###rpart import
library(rpart)
library(rpart.plot)


#Reminder: Price column set to NULL after creating price_update class column
model_tree <- rpart(price_update ~.,data = training)
###summary
summary(model_tree)
```
 
### Full Tree Plot
```{r Full Tree}

###full tree plot
set.seed(757900)
tree_full <- rpart(formula=price_update~.,data=training,
                control=rpart.control(cp=0.0010,minsplit=50,xval=10))
rpart.plot(tree_full, box.palette="RdBu", shadow.col="gray", nn=TRUE)

```


Control Parameter Check - Full Tree
```{r print tree_full}
printcp(tree_full)
```

### Pruned Tree Plot
```{r pruned tree plot}
###pruned tree plot
set.seed(757900)
tree_prune <- rpart(formula=price_update~.,data=training,
                control=rpart.control(cp=0.0020,minsplit=5,xval=10))
rpart.plot(tree_prune, box.palette="RdBu", shadow.col="gray", nn=TRUE)
```


```{r Test Data Full vs. Prune Tree}
###preidct using tree on test data
testing$full_tee_pred <- predict(tree_full,testing, type= "class")
testing$prune_tree_pred <- predict(tree_prune,testing, type ="class")

```

### Confusion Matix Full Tree
```{r full tree confusion matrix}
##confusion matrix for the full tree
confusionMatrix(testing$full_tee_pred,testing$price_update)
```

### Confusion Matix Pruned Tree

```{r prurned tree confusion matrix}
##confusion matrix for the pruned tree
confusionMatrix(testing$prune_tree_pred,testing$price_update)
```

## Random Forest Model
```{r Random Forest}
library(dplyr)
training=training %>% mutate_if(is.character, as.factor)
library(randomForest)
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

RF_model <- randomForest(price_update~., data = training, trControl=trctrl, importance=TRUE,proximity=TRUE,ntrees = 100)
```
redict on the random forest model and  caclulate the accuracy of the model
```{r random forest train, test}
testing=testing %>% mutate_if(is.character, as.factor)

pred_test <- predict(RF_model, testing, type="class")
pred_train <- predict(RF_model, training, type = "class")

confusionMatrix(pred_test, testing$price_update)
confusionMatrix(pred_train, training$price_update)


```